Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-11-11 paper/main.tex
q

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\usepackage{mathrsfs} % math script fonts
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
% *** SPECIALIZED LIST PACKAGES ***
\usepackage{xcolor}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{subfig} %ieee does not like subfigure
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?
\usepackage[switch]{lineno}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   framesep=3mm} 
% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{tabulary}
% IEEEtran contains the IEEEeqnarray family of commands

@@

@@

\begin{definition} 
  \label{def:category:F} 
  The fiber category $\mathcal{\dfiberc}$ is a product category\cite{maclaneCategoriesWorkingMathematician2013,nlab:product_category} of zero or more arbitrary categories. In the two category case, the category $\mathcal{\dfiberc}$ consists of 
  \begin{itemize}
    \item \textit{objects} ordered pairs $(x,y)$,  $x$ is an object of $\mathcal{X}$ and  $y$ is an object of $\mathcal{Y}$
    \item \textit{morphisms} ordered pairs $(f:x\rightarrow x^{\prime}, g:y\rightarrow y^{\prime})$
    \end{itemize}
The coomposition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
\end{definition}

@@

The domain of the artist function is the space of all possible graphics. As with the domain $\sheaf_{E}$, the grapic space is modeled with fiber bundle equivalent categories:
\begin{table}[!h]
  \centering
  \label{tab:artist:graphic}
  \begin{tabular}{l l l}
    fiber bundle & data & graphic \\
    \textcolor{base}{base space} & \mathcal{\dbasec} & \mathcal{\gbasec}\\
    \textcolor{total}{total space} & \mathcal{\dtotalc} & \mathcal{\gtotalc}\\
    \textcolor{fiber}{fiber space} & \mathcal{\dfiberc} & \mathcal{\gfiberc}\\
    \pi^{-1} & \sheaf_{\mathcal{\dtotalc}} & \sheaf_{\mathcal{\gtotalc}}\\
  \end{tabular}
\end{table}

@@

\begin{equation}
  \label{eq:artist:construction:all}
  \begin{tikzcd}
    \dtotalc \arrow[rr, "\vchannel"] \arrow[rdd, "\pi"'] &         & \dtotalc \arrow[ldd, "\pi"] & \mathcal{\vtotalc}                                                        & Im_{\vmarkc} \subset \gtotalc                     \\
  &         &                             & {} \arrow[r, "\vmarkc"]                                                   & {}                                \\
   & \dbasec &                             & \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] \arrow[r, "\vindexc^*"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vmarkc}"']
  \end{tikzcd}
\end{equation}
which are the data to visual variable map $\vchannel$ at the bundle level and the visual variable to graphic map $\vmark$ at the sheaf level. Usage of the bundle directly in an equation or diagram - $\dotal, \vtotal, \gtotal$ - denotes that the function can be evaluated pointwise over $\dbasepoint \in \dbase$. Use of the categorical form -- $\mathcal{\dtotalc}, \mathcal{\vtotalc}, \mathcal{\gtotalc}$ -- denotes that the function is evaluated over an openset object, meaning that the function needs knowledge of both a value over a point and some information about neighboring values.

\subsection{Data Domain}
We model data as sections of a fiber bundle $(\dtotal, \pi, \dbase, \dfiber)$. We encode the continuity of the data as the \textcolor{base}{base space} $\dbase$. \note{does this go in intro?} The \textcolor{fiber}{fiber space} \dfiber\ is the space of all possible values of the data \autoref{eq:related-work:fiber}. We model the data as the section \dsection because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase to the space of possible data values \dfiber. Spivak's notation, as discussed in \autoref{sec:related-work:equivariance}, also allows for associating elements in a section with the field it comes from. This allows for field based selection of values, while inclusion allows for continuity (index) based selections.

@@

\subsubsection{Visual Bundle \vtotal}
\begin{equation}
  \begin{tikzcd}[ampersand replacement=\&]
      \vfiber \arrow[r, hook] \& \vtotal \arrow[d, "\pi"'] \\
                        \& \dbase \arrow[u, "\vsection"', bend right]
  \end{tikzcd}
\end{equation}

\subsubsection{Graphic to Data: \vindex}
\begin{equation}
  \label{eq:artist:construction:xi}
  \begin{tikzcd}
      \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
      \dbase                   & \gbase \arrow[l, "\vindex"']
  \end{tikzcd}
\end{equation}

\begin{equation}
  \label{eq:artist:construction:xi:multi_base}
  \begin{tikzcd}
    \dbase_{c} \arrow[d, hook] \arrow[r, hook] & \dbase_{b}                 & \gbase_b \arrow[l, "\vindex"]                                                \\
    \dbase_{a}                                 & \gbase_a \arrow[l, "\vindex"'] & \gbase_c \arrow[l, hook'] \arrow[u, hook] \arrow[llu, "\vindex" description]
  \end{tikzcd}
\end{equation}

Can be implemented in V

\begin{equation}
  \label{eq:artist:construction:xi:v}
  \begin{tikzcd}
  \textcolor{color=gray!50}{\dtotal}\arrow[rd, "\pi"', color=gray!50] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', dashed] & \textcolor{color=gray!50}{\gtotal} \arrow[ld, "\pi", color=gray!50] \\
    & \dbase                   & \gbase \arrow[l, "\vindex"']                                                &                          
    \end{tikzcd}
\end{equation}

\subsubsection{Data to Visual Encodings: \vchannel} 
\begin{equation}
  \label{eq:artist:construction:nu}
\begin{tikzcd}
  \dtotal \arrow[rd, "\pi"'] \arrow[r, "\vchannel"] & \vtotal \arrow[d, "\pi"] & \textcolor{fade}{\vindex^*\vtotal} \arrow[d, "\vindex^*\pi"', color=fade] \arrow[l, "\vindex^*", color=fade] & \textcolor{fade}{\gtotal} \arrow[ld, "\pi", color=fade] \\
& \dbase                   & \textcolor{gray!50}{\gbase} \arrow[l, "\vindex"', color=gray!50]                                       &                          
  \end{tikzcd}
\end{equation}

Visual encoding functions \vindex\ are functors because they are expected to preserve structure on the data and visual side, as discussed in \autoref{sec:related-work:equivariance}.

 This constraint is expressed in our construction of \vchannel

\begin{figure}[!h]
  \centering
  \subfloat[Artists with shared $\mu_i$ renderered correctly]{\includegraphics[width=1.2in]{partial_fixed.png}%
  \label{fig_first_case}}
  %\hfil
  \subfloat[Artists without shared $\mu_i$]{\includegraphics[width=1.2in]{partial_invalid.png}%
  \label{fig_second_case}}
  \caption{Simulation results for the network.}
  \label{fig_sim}
  \end{figure}

\begin{equation}
\label{eq:artist:construction:nu:fiber_map}
\begin{tikzcd}[column sep=.5em]
  \dfiber \arrow[rr, "\vchannel"] &                                                         & \vfiber \\
                                  & \dbase \arrow[lu, "\dsection"] \arrow[ru, "\vsection"'] &        
  \end{tikzcd}
\end{equation}

which means equivariance of 
\begin{equation}
\label{eq:artist:construction:nu:equivariance}
\begin{tikzcd}
  \dsection \arrow[d, "\varphi"'] \arrow[r, "\vchannel"] & \vsection \arrow[d, "\vchannel_{\varphi}"] \\
  \dsection^{\prime} \arrow[r, "\vchannel"]                       & \vsection^{\prime}                                 
  \end{tikzcd}
\end{equation}

@@

\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}

@@

\subsubsection{Visual to Graphic: \vmark} 
\begin{equation}
  \label{eq:artist:construction:q}
  \begin{tikzcd}
    \textcolor{fade}{\dtotal} \arrow[rd, "\pi"', color=fade] \arrow[r, "\vchannel", color=fade] & \textcolor{gray!50}{\vtotal} \arrow[d, "\pi", color=fade] & \vindex^*\vtotal \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"', color=fade] \arrow[r, "\vmark"] & \gtotal \arrow[ld, "\pi"] \\
    & \textcolor{fade}{\dbase} & \gbase \arrow[l, "\vindex"', color=fade]                                            &                          
    \end{tikzcd}
\end{equation}

\begin{equation}
  \label{eq:artist:construction:q:xi}
  \begin{tikzcd}
    \vtotalc \arrow[rr, "\vmarkdc", bend left] & \vtotalpullc \arrow[l, "\vindexc^*"] \arrow[r, "\vmarkc"'] & \gtotalc
  \end{tikzcd}
\end{equation}

\begin{equation}
  \label{eq:artist:construction:q:fiber}
  \begin{tikzcd}[column sep=.5em]
    \vfiber \arrow[rr, "\vmark"] &                                                         & \gfiber \\
    & \gbase \arrow[ru, "\gsection"'] \arrow[lu, "\vsection"] &        
    \end{tikzcd}
\end{equation}
Visual encodings to something like marks

\begin{equation}
  \label{eq:artist:construction:q:equivariance}
  \begin{tikzcd}
    \vsection \arrow[d, "\vchannel_\varphi"'] \arrow[r, "\vmark"] & \gsection \arrow[d, "\vmark_\varphi"] \\
    \vsection^{\prime} \arrow[r, "\vmark"]                                 & \gsection^{\prime}                            
    \end{tikzcd}
\end{equation}

@@

\begin{equation}
  \label{eq:artist:construction:q:multifiber}
\begin{tikzcd}
  \dfiberc_a \arrow[rr, "\vchannelc"] \arrow[rd, "proj_a" description]          &            & \vfiberc_a \arrow[rr, "\vmarkd"]         &  & Im_{\vmarkdc} \subset \gtotalc \arrow[llld, "g_a" description, dashed]                  \\
 & \dfiberc_c &                                          &  &\\
  \dfiberc_b \arrow[rr, "\vchannelc^{\prime}"] \arrow[ru, "proj_b" description] &            & \vfiberc_b \arrow[rr, "\vmarkd^{\prime}"] &  & Im_{\vmarkdc^{\prime}}\subset \gtotalc^{\prime} \arrow[lllu, "g_b" description, dashed]
  \end{tikzcd}
\end{equation}

\section{Case Study}
\label{sec:case-study}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{path_of_q.png}
    \caption{\note{add in xi!}}
  \label{fig:api}
\end{figure}
We implement the \textbf{arrows} in \autoref{fig:api}. \mintinline{python}{axesArtist} is a parent artist that acts as a screen. This allows for the composition described in                                           \autoref{sec:artist:union}

>>>>

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\usepackage{mathrsfs} % math script fonts
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
% *** SPECIALIZED LIST PACKAGES ***
\usepackage{xcolor}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{subfig} %ieee does not like subfigure
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   framesep=3mm} 
% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{tabulary}
% IEEEtran contains the IEEEeqnarray family of commands

@@

\usepackage[switch]{lineno}
\linenumbers

@@

\begin{definition} 
  \label{def:category:F} 
  The fiber category $\mathcal{\dfiberc}$ is a product category\cite{maclaneCategoriesWorkingMathematician2013,nlab:product_category} of zero or more arbitrary categories. In the two category case, the category $\mathcal{\dfiberc}$ consists of 
  \begin{itemize}
    \item \textit{objects} ordered pairs $(x,y)$,  $x$ is an object of $\mathcal{X}$ and  $y$ is an object of $\mathcal{Y}$
    \item \textit{morphisms} ordered pairs $(f:x\rightarrow x^{\prime}, g:y\rightarrow y^{\prime})$
    \end{itemize}
The composition of morphisms is defined component wise by composition in $\mathcal{X}$ and $\mathcal{Y}$.ss
\end{definition}
The fiber category $\mathcal{\dfiber}$ can encode a large variety of collections of data fields because it is a product category of arbitrary categories. For example, a lists of strings is an instance of an object in \textbf{Set}, networks are an instance of \textbf{Graph}, and images are vector spaces which are a specific type of topological space \textbf{Top}.  The morphisms $Hom_{\mathcal{\dfiber}}$ are functions from the data to itself, for example the binary operations, group actions, and measurement scales discussed in \autoref{sec:related-work:equivariance}. These functions could also be monoid actions\cite{barrCategoryTheoryComputing}, which provide a way of applying partial order relations to data\cite{fongInvitationAppliedCategory2019}, such as to build multi-ranked indicators\cite{bruggemannRankingPrioritizationMultiindicator2011}.

\note{note sure this note applies anymore: We note that a one object category is a monoid \cite{nlab:monoid,barrCategoryTheoryComputing}, which is an algebraic structure that lends itself to computer library design because they are an abstraction of function composition \cite{yorgeyMonoidsThemeVariations}.}

@@

The domain of the artist function is the space of all possible graphics. As with the domain $\sheaf_{E}$, the graphic space is modeled with fiber bundle equivalent categories:
\begin{table}[!h]
  \centering
  \label{tab:artist:graphic}
  \begin{tabular}{l l l}
    fiber bundle & data & graphic \\
    \textcolor{base}{base space} & \mathcal{\dbasec} & \mathcal{\gbasec}\\
    \textcolor{total}{total space} & \mathcal{\dtotalc} & \mathcal{\gtotalc}\\
    \textcolor{fiber}{fiber space} & \mathcal{\dfiberc} & \mathcal{\gfiberc}\\
    \pi^{-1} & \sheaf_{\mathcal{\dtotalc}} & \sheaf_{\mathcal{\gtotalc}}\\
  \end{tabular}
\end{table}

@@

\begin{equation}
  \label{eq:artist:construction:all}
  \begin{tikzcd}
    \dtotalc \arrow[r, "\vindexc"] \arrow[dd, "\pi"'] & \vtotalc \arrow[ldd, "\pi"] & \mathcal{\vtotalc}                                                        & \gtotalc                                           \\
    &                             & {} \arrow[r, "\vmarkc"]                                                   & {}                                                 \\
    \dbasec                                           &                             & \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] \arrow[r, "\vindexc^*"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vartist}"']
    \end{tikzcd}
\end{equation}
which are the data to visual variable map $\vchannel$ at the bundle level and the visual variable to graphic map $\vmark$ at the sheaf level. Usage of the bundle directly in an equation or diagram - $\dotal, \vtotal, \gtotal$ - denotes that the function can be evaluated pointwise over $\dbasepoint \in \dbase$. Use of the categorical form -- $\mathcal{\dtotalc}, \mathcal{\vtotalc}, \mathcal{\gtotalc}$ -- denotes that the function is evaluated over an openset object, meaning that the function needs knowledge of both a value over a point and some information about neighboring values.

\subsection{Data Domain}
We model data as sections of a fiber bundle $(\dtotal, \pi, \dbase, \dfiber)$. We encode the continuity of the data as the \textcolor{base}{base space} $\dbase$. We model the data as the section \dsection\ because, as described in \autoref{sec:related-work:fiber-bundles}, the section is the map from the indexing space \dbase\ to the space of possible data values \dfiber. We adopt Spivak's formulation of the fiber as a (column name, data domain) simple schema \cite{spivakSIMPLICIALDATABASES,spivakDatabasesAreCategories2010}. Spivak formally maps column names and field types to the set of values associated with the field type, for example $\mathbb{R}$ for a \texttt{float} column named \textit{temperature}. This allows for field based selection of values, while inclusion allows for continuity (index) based selections.

@@

\subsubsection{Visual Bundle \vtotal}
\begin{equation}
  \begin{tikzcd}[ampersand replacement=\&]
      \vfiberc \arrow[r, hook] \& \vtotalc \arrow[d, "\pi"'] \\
                        \& \dbasec \arrow[u, "\vsectionc"', bend right]
  \end{tikzcd}
\end{equation}

\subsubsection{Data to Visual Encodings: \vchannel} 
\note{maybe figure out how to put these side by side nicely}

\begin{equation}
  \label{eq:artist:construction:nu}
  \begin{tikzcd}
    \dtotalc \arrow[d, "\pi"'] \arrow[r, "\vchannelc"] & \vtotalc \arrow[ld, "\pi"] &  & \dfiberc \arrow[r, "\vchannelc"]                          & \vfiberc \\
    \dbasec                                            &                            &  & \dbasec \arrow[u, "\dsectionc"] \arrow[ru, "\vsectionc"'] &         
    \end{tikzcd}
\end{equation}
Since the functor \vindex\ is bundle wise, it can be evaluated at the bundle at a point, which is the fiber space; therefore \vindex\ is a functors from the product category $\mathcal{\dfiber}$ to the product categrory $\mathcal{\vfiber}$. This constraint is expressed in our construction of \vchannel

which means equivariance of 
\begin{equation}
\label{eq:artist:construction:nu:equivariance}
\begin{tikzcd}
  \dsectionc \arrow[d, "\dfuncc"'] \arrow[r, "\vchannelc"] & \vsectionc \arrow[d, "\vchannelc_{\dfuncc}"] \\
  \dsectionc^{\prime} \arrow[r, "\vchannelc"]                       & \vsectionc^{\prime}                                 
  \end{tikzcd}
\end{equation}

Since the functor $\nu$ acts on product categories, it can be decomposed into $nu_i$ component functors that act on corresponding sections $\dsection_i$ of the fiber $\dfiber_i$.

\begin{equation}
  \label{eq:math:artist:nu}
  \{\vchannelc_{0}, \ldots, \vchannelc_{n}\}: \{\dsectionc_{0}, \ldots, \dsectionc_{n}\} \mapsto \{\vsectionc_{0}, \ldots, \vsectionc_{n}\}
\end{equation}

One example of this is

\begin{figure}[!h]
  \centering
  \subfloat[\note{insert title}]{\includegraphics[width=1.2in]{partial_fixed.png}%
  \label{fig_first_case}}
  %\hfill
  \subfloat[\note{insert title}]{\includegraphics[width=1.2in]{partial_invalid.png}%
  \label{fig_second_case}}
  \caption{is a weird nu, colors are a but much}
  \label{fig_sim}
\end{figure}

@@

@@

\subsubsection{Visual to Graphic: \vmark}

\begin{equation}
  \label{eq:artist:construction:q}
  \begin{tikzcd}
    \mathcal{\vtotalc}                                & \gtotalc                                           \\
    {} \arrow[r, "\vmarkc"]                           & {}                                                 \\
    \mathcal{\dbasec} \arrow[uu, "\sheaf_{\vtotalc}"] & \mathcal{\gbasec} \arrow[uu, "\sheaf_{\vartist}"']
    \end{tikzcd}
\end{equation}

@@

\begin{equation} 
  \label{eq:artist:construction:latent_vars}
\begin{tikzcd}
  \mathcal{\vfiberc_a} \arrow[rrr, "\vmarkc_a"] \arrow[rd, "p_c"]          &                      &  & \sheafc_{\vartist} \arrow[lld, "l_c", dashed]                     \\
                                                                           & \mathcal{\vfiberc_c} &  &                                                                   \\
  \mathcal{\vfiberc_b} \arrow[rrr, "\vmarkc_b"] \arrow[ru, "p^{\prime}_c"] &                      &  & \sheafc_{\vartistc^{\prime}} \arrow[llu, "l^{\prime}_c"', dashed]
\end{tikzcd}
\end{equation}

\subsubsection{Graphic to Data: \vindex}
\begin{equation}
  \label{eq:artist:construction:xi}
  \begin{tikzcd}
    \textcolor{fade}{\dtotal} \arrow[rd, "\pi"', color=fade] &         & \textcolor{fade}{\vtotal} \arrow[ld, "\pi", color=fade] &  & \textcolor{fade}{\gtotal} \arrow[d, "\pi", color=fade] \\
  & \dbasec &                                             &  & \gbasec \arrow[lll, "\vindexc"']          
    \end{tikzcd}
\end{equation}

By \note{homotopy lifting?}
\begin{equation}
  \label{eq:artist:construction:xi:lifted}
  \begin{tikzcd}
    & \opensetc_j \arrow[d, "\iota", hook] \\
\dbasec \arrow[ru, "\xi^*", dashed] & \gbasec \arrow[l, "\xi"]            
\end{tikzcd}
\end{equation}

\begin{equation}
  \label{eq:artist:construction:xi:multi_base}
  \begin{tikzcd}
    \dbasec_{c} \arrow[d, hook] \arrow[r, hook] & \dbasec_{b}                 & \gbasec_b \arrow[l, "\vindexc"]                                                \\
    \dbasec_{a}                                 & \gbasec_a \arrow[l, "\vindexc"'] & \gbasec_c \arrow[l, hook'] \arrow[u, hook] \arrow[llu, "\vindexc" description]
  \end{tikzcd}
\end{equation}

\note{talk about xi even though is not explicitely implemented}
The mapping between graphic and data space expresses how...
This is what allows the artist to generate graphics where the subset of data on view is dynamically updated, such as pan and zoom navigation\cite{NekrasovskiEvaluationPanZoom2006} and sliding windows on streaming data\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

\section{Case Study}
\label{sec:case-study}
\begin{figure}[h!]
  \includegraphics[width=\columnwidth]{path_of_q.png}
    \caption{\vchannel \& \mark are what we're implementing}
  \label{fig:api}
\end{figure}
We implement the \textbf{arrows} in \autoref{fig:api}. \mintinline{python}{axesArtist} is a parent artist that acts as a screen. This allows for the composition described in                                           \autoref{sec:artist:union}
